Contextual recruitment of selective attention can be updated via changes in task-relevance

### 2019-10-08 manuscript_pretty.tex
so close

\abstract{
Evidence across a wide variety of attention paradigms shows that environmental cues can trigger adjustments to ongoing priorities for attending to relevant and irrelevant information. This context-specific control over attention suggests that cognitive control can be both automatic and flexible. For instance, in selective attention tasks, congruency effects are larger for items that appear in a context associated with infrequent conflict than in a context associated with frequent conflict. Since the to-be-presented context cannot be predicted or prepared for in advance, attention is assumed to be rapidly updated on-the-fly, triggered by the currently presented context. Context-specific control exemplifies how learning and memory processes can influence attention to enable cognitive flexibility. However, what determines the use of previously learned associations still remains unclear. In the current study, we examined whether task-relevance would influence the learning and use of context cues in a flanker task. Using a secondary counting task, context dimensions associated with differing levels of conflict were made task-relevant or -irrelevant across the experiment. In short, we found that making new contextual information task-relevant caused participants to ignore a previously learned context-attention association and adopt a new context-specific control strategy; all without changing the experimental stimuli. This result suggests that task-relevance is a key determinant of context-specific control.

@@

Selective attention is commonly investigated using interference paradigms like the Stroop (1935) and flanker (Eriksen \& Eriksen, 1974) tasks, where participants identify a target while ignoring a response-congruent or -incongruent distractor. Performance is typically better on congruent versus incongruent trials and the difference---the congruency effect---taken as an index of attentional priorities. Large congruency effects are thought to reflect ineffective filtering of the distracting stimuli whereas small congruency effects are thought to reflect effective filtering. By probing factors that systematically alter congruency effects, we can then make inferences about processes that control attentional filtering. For example, manipulating the frequency of conflict via the proportion of congruent versus incongruent trials has shown to influence the size of the congruency effect. Typically, a high proportion congruent experiment produces large congruency effects, whereas a low proportion congruent experiment produces small congruency effects (Logan \& Zbrodoff, 1979; Lowe \& Mitterer, 1982; West \& Baylis, 1998). This result is usually explained as strategic control, where participants increase attentional control under high-conflict demands and relax attentional control under low-conflict demands (Logan, 1980; Logan \& Zbrodoff, 1979; Logan, Zbrodoff, \& Williamson, 1984; Lowe \& Mitterer, 1982). Recent work however, has demonstrated that attentional control is not only adjusted by top-down regulation, but can also be triggered automatically by environmental cues (Brosowsky \& Crump, 2018; Bugg \& Crump, 2012; Egner, 2014; Fischer \& Dreisbach, 2015; King, Korb, \& Egner, 2012; Mayr \& Bryck, 2007).

@@

The current study addresses two related issues: First, how is attentional control resolved when multiple contextual features could be used as cues? And second, what determines the expression of previously learned attentional associations?

Real-world environments are complex and contain many potential context cues. For instance, consider someone writing in a coffee shop. The coffee shop is a rich environment with numerous context cues (e.g., pictures on the wall, chairs, coffee, laptop), each of which we have different experiences with and afford different learned associations. Although a laptop could cue attentional processing helpful for writing, other features of the coffee shop might cue attentional processing related to other experiences like socializing and people-watching. Given that all of these cues are available to set attentional priorities, what determines which associative relationships will be used to guide attention? On the one hand, context-based retrieval of attentional priorities might be completely obligatory: the mere presence of learned cues might trigger attention in a ballistic manner (e.g., Schneider \& Shiffrin, 1977; Shiffrin \& Schneider, 1977). A writer's attention in this case would be at the mercy of the environment and attention is guided by the strongest association. On the other hand, our current task-goals might constrain the expression of previously learned associations. The writer's current goals may allow the laptop to take priority over other environmental cues that are irrelevant to the task at hand.

Historically, perspectives of automaticity and attentional control have emphasized the role of practice and likened the learning process to the development of automaticity in skill-learning (Logan, 1988, 1992; Schneider \& Shiffrin, 1977; Shiffrin \& Schneider, 1977). Under this view, forming an association requires many repeated experiences pairing a cue with a set of attentional priorities. Once formed however, it becomes habitual or hard-wired and the mere presence of the cue is sufficient to automatically adjust attention. Learned associations are therefore ballistic, difficult to suppress, and difficult to unlearn (e.g., Schneider \& Shiffrin, 1977; Posner \& Snyder, 1975a, 1975b; Shiffrin \& Schneider, 1977; Treisman, 1960, 1964, 1969). This view is reflected in modern work on contextual cuing where it is assumed that learned associations require practice (e.g., Lehle \& Hübner, 2008; Crump, 2016), form without awareness or intention, and are triggered automatically by the mere presence of the cue (Crump et al., 2006; Crump \& Logan, 2010; Crump, Vaquero, \& Milliken, 2008; King et al., 2012; Reuss, Desender, Kiesel, \& Kunde, 2014; though, see Brosowsky \& Crump, 2016).

Memory-based perspectives however, take a more nuanced view (Abrahamse et al., 2016; Brosowsky \& Crump, 2018; Crump, 2016; Egner, 2014; Norman, 1969, 1976). Here, practice is only important to the extent that it improves memory retrieval. Learning an association is the process of encoding experiences where attentional priorities and cues co-occur. The expression of previously learned associations however, is dictated by memory retrieval---a similarity-based, cue-driven process (e.g., Jacoby, 1978; Hintzman, 1984, 1986, 1988). Presumably, the features of our immediate environment automatically cue memory and reinstate the retrieved attentional priorites (Logan, 1988, 1992). If a few prior experiences are easily retrievable (i.e., memories are distinct or very similar to the retrieval cue), they could support the reinstatement of a learned attentional association despite very little practice (e.g., Brosowsky \& Crump, 2018). Conversely, if prior experiences are difficult to retrieve (i.e., memories are indistinct or low in similarity to the retrieval cue), then learned associations may not be expressed, even after much practice. To the extent that you could manipulate memory retrieval, you should als be able manipulate the expression of learned associations.

Here, we examine whether task-relevance influences the retrieval of previously used attentional priorities. We hypothesized that the cue-driven, memory-retrieval process might be constrained to the features of our environment that are relevant to our on-going task and goals. If this is the case, then a previously learned association should be easily ignored if that feature was made irrelevant to the task because that feature would no longer contribute to memory-retrieval. Furthermore, it would suggest that the mere presence of a cue is not sufficient for automatically directing attention, but instead requires some task-integration.

Recent work lends some credence to the possibility that task-relevance is necessary for establishing new context-attention associations. Crump, Vaquero, \& Milliken (2008) used a Stroop prime-probe task where the distractor word is presented first followed by a color patch probe. Critically, they used the shape of the color patch probe, either a square or a circle, as the contextual dimension; one associated with a high frequency of conflict, the other with a low frequency. In one experiment participants were made aware of the shape-proportion contingencies but given no additional instructions. In the second, participants were given a secondary task to count the number of trials that contained a square. The first experiment failed to find evidence for context-specificity, whereas the second did. Although not testing task-relevance directly, Cañadas, Rodríguez-Bailón, Milliken, and Lupiáñez (2013) found similar effects. In this study they used images of male and female faces as context cues in a flanker task. When given instructions to think of the faces as members of the gender categories they found CSPC effects, but when given instructions to think of the faces as individuals they did not. These results suggest that making a context dimension relevant to the on-going task is important for establishing new associations and producing CSPC effects. However, the extent to which task-relevance influences the expression of previously learned associations is still unknown.

@@

Prior work suggests that the task-relevance of contextual information could be important for establishing new associations between context cues and attentional priorities. Simply put, CSPC effects tend to emerge when context cues are somehow made relevant to the task, but do not when made irrelevant (Cañadas et al., 2013; Crump et al., 2006, 2008). However, what determines the learning and expression of context-attention associations remains unclear. In the current study, we examined whether manipulating task-relevance would allow participants to ignore a previously learned association and adopt a new context-specific control strategy.

@@

The critical measure of interest is the congruency effect produced in the frequency unbiased context. If, on the one hand, task-relevance has no impact on the use of context-attention associations we would expect no differences between the congruency effects when the high-conflict context dimension is made task-relevant or the low-conflict context dimension is made task-relevant. This could occur because all three contexts are treated as individual contexts throughout the whole experiment (e.g., Cañadas et al., 2013). Or it could occur because associations formed in the first block interfere with learning new associations in the second (e.g., Brosowsky \& Crump, 2016). On other hand, task-relevance may dictate the expression of previously learned associations. In this case, we would expect smaller congruency effects when the high-conflict context dimension is made task-relevant as compared to when the low-conflict context dimension is made task-relevant. This would demonstrate that participants were able to ignore a previously learned association and adopt a new context-specific control strategy.

@@

Experiment 1 used a 2x2x3 mixed design with task-relevant context (0\% and 100\%) and unbiased-item congruency (congruent and incongruent) as within-subject factors, and context-type (object, social, and social/non-repeating) as the between-subjects factor.

@@

@@

\begin{table}[htbp]
\caption{Reaction times and error rates from Experiment 1.}
\label{TR_table}
\centering
\begin{tabular}{lllcccc}
\toprule
 & & & \multicolumn{2}{c}{Congruent} & \multicolumn{2}{c}{Incongruent} \\
\cmidrule(rl){4-5}
\cmidrule(rl){6-7}
\multicolumn{1}{c}{Task-Relevant Context} & \multicolumn{1}{c}{Condition} & \multicolumn{1}{c}{PC} & \multicolumn{1}{c}{RT} & \multicolumn{1}{c}{ER} & \multicolumn{1}{c}{RT} & \multicolumn{1}{c}{ER}  \\
\midrule
\multirow{9}{*}{100\% PC} & \multirow{3}{*}{Object} & \multicolumn{1}{l}{0\%} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{693 (26)} & \multicolumn{1}{l}{4.17 (0.73)} \\
& & \multicolumn{1}{l}{50\%} & \multicolumn{1}{l}{585 (23)} & \multicolumn{1}{l}{0.74 (0.27)} & \multicolumn{1}{l}{675 (28)} & \multicolumn{1}{l}{4.81 (0.82)} \\
& & \multicolumn{1}{l}{100\%} & \multicolumn{1}{l}{587 (25)} & \multicolumn{1}{l}{0.69 (0.23)} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} \\
\cmidrule(rl){2-7}
& \multirow{3}{*}{Social} & \multicolumn{1}{l}{0\%} & \multicolumn{1}{c}{-} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{745 (25)} & \multicolumn{1}{l}{5.86 (0.77)} \\
& & \multicolumn{1}{l}{50\%} & \multicolumn{1}{l}{612 (21)} & \multicolumn{1}{l}{0.39 (0.19)} & \multicolumn{1}{l}{722 (26)} & \multicolumn{1}{l}{6.78 (1.04)} \\
& & \multicolumn{1}{l}{100\%} & \multicolumn{1}{l}{616 (20)} & \multicolumn{1}{l}{0.63 (0.2)} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} \\
\cmidrule(rl){2-7}
& \multirow{3}{*}{Social (NR)} & \multicolumn{1}{l}{0\%} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{813 (25)} & \multicolumn{1}{l}{5.04 (0.74)} \\
& & \multicolumn{1}{l}{50\%} & \multicolumn{1}{l}{689 (23)} & \multicolumn{1}{l}{0.19 (0.14)} & \multicolumn{1}{l}{793 (27)} & \multicolumn{1}{l}{5.72 (1.06)} \\
& & \multicolumn{1}{l}{100\%} & \multicolumn{1}{l}{689 (25)} & \multicolumn{1}{l}{0.48 (0.17)} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} \\
\midrule
\multirow{9}{*}{0\% PC} & \multirow{3}{*}{Object} & \multicolumn{1}{l}{0\%} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{649 (16)} & \multicolumn{1}{l}{3.84 (0.6)} \\
& & \multicolumn{1}{l}{50\%} & \multicolumn{1}{l}{587 (17)} & \multicolumn{1}{l}{0.56 (0.21)} & \multicolumn{1}{l}{657 (17)} & \multicolumn{1}{l}{4.17 (1.02)} \\
& & \multicolumn{1}{l}{100\%} & \multicolumn{1}{l}{634 (18)} & \multicolumn{1}{l}{2.69 (0.58)} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} \\
\cmidrule(rl){2-7}
& \multirow{3}{*}{Social} & \multicolumn{1}{l}{0\%} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{722 (23)} & \multicolumn{1}{l}{5.77 (0.92)} \\
& & \multicolumn{1}{l}{50\%} & \multicolumn{1}{l}{633 (20)} & \multicolumn{1}{l}{0.78 (0.5)} & \multicolumn{1}{l}{722 (25)} & \multicolumn{1}{l}{4.75 (0.9)} \\
& & \multicolumn{1}{l}{100\%} & \multicolumn{1}{l}{673 (21)} & \multicolumn{1}{l}{1.84 (0.47)} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} \\
\cmidrule(rl){2-7}
& \multirow{3}{*}{Social (NR)} & \multicolumn{1}{l}{0\%} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{807 (26)} & \multicolumn{1}{l}{4.55 (0.8)} \\
& & \multicolumn{1}{l}{50\%} & \multicolumn{1}{l}{735 (28)} & \multicolumn{1}{l}{0.87 (0.36)} & \multicolumn{1}{l}{818 (30)} & \multicolumn{1}{l}{4.26 (0.8)} \\
& & \multicolumn{1}{l}{100\%} & \multicolumn{1}{l}{779 (34)} & \multicolumn{1}{l}{2.03 (0.4)} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} \\
 & & & & & & \\
\bottomrule
\multicolumn{7}{l}{\textit{Note}: RT = Reaction Time (ms);  ER = Error Rates (\%); PC = Proportion Congruent;} \\
\multicolumn{7}{l}{NR = Non-Repeating; Standard Errors are presented in parantheses.} \\
\end{tabular}%
\end{table}

@@

In Experiment 1, we found that manipulating the task-relevance of a shared contextual feature was associated in predictable changes in context-specific proportion congruency effects for frequency unbiased context items; Namely, larger congruency effects when the context was associated with low conflict than when the context was associated with high conflict. This suggests that participants learned the association between context and conflict-level in the first block (despite multiple, competing cues), and subsequently ignored that association in the second block to learn a new association. Task-relevance then, may not just be important for establishing new associations (e.g., Crump \& Milliken, 2009) but also important for determining which learned associations are expressed in the present moment.

@@

Experiment 2, therefore, serves as a control experiment to test these two assumptions directly. Experiment 2 is identical to Experiment 1 with two important changes: First, we removed the counting task, giving no explicit instructions about how to use the context cues; We only informed participants that images would be present throughout the task. This allowed us to determine if there was any systematic effect of biased items on the unbiased items absent the counting task. Second, we altered the proportion congruency of the biased item sets from 100\%/0\% to 75\%/25\%. This allowed us to measure the congruency effect for the biased item sets to determine whether participants learned about the associations between proportion congruency and the context cues without the counting task (e.g., Crump \& Milliken, 2009). If this were the case, we would expect larger congruency effects for the 75\% proportion congruent items as compared to the 25\% proportion congruent items.

@@

@@

In the current study, we have been examining the importance of task-relevance for the contextual recruitment of selective attention. In Experiment 1, we found that task-relevance allowed participants to exploit context cues in the presence of a competing, contradictory cue and subsequently ignore a previously learned association in favor of a different context cue. In Experiment 2, we conceptually replicated previous work showing that task-relevance was necessary for participants to learn about the biased proportion congruency associations (e.g., Crump \& Milliken, 2009) and ruled out possible confounding effects of the biased items.

Experiment 3 tests the boundaries of the task-relevance effect found in Experiment 1. The evidence thus far suggests that context-cues, rather than being irrelevant to the task (as their namesake would suggest), need to be integrated into the task-representation to guide attentional control. There is some evidence that instructions alone can be sufficient to produce list-wide proportion congruency effects, though not item-specific proportion congruency effects (e.g., Entel et al., 2014; though, see Desender, 2018) and not context-specific proportion congruency effects (e.g., ). Explicit instructions however, are another way of making context-cues relevant to the on-going task and could serve the same purpose as the counting task---although, instructions are likely a much weaker manipulation than the counting task. In Experiment 3, we test whether participants are able to explicitly use the context-cues to guide attentional control through an instruction manipulation.

@@

@@

>>>>

\abstract{
Evidence across a wide variety of attention paradigms shows that environmental cues can trigger adjustments to ongoing priorities for attending to relevant and irrelevant information. This context-specific control over attention suggests that cognitive control can be both automatic and flexible. For instance, in selective attention tasks, congruency effects are larger for items that appear in a context associated with infrequent conflict than in a context associated with frequent conflict. Since the to-be-presented context cannot be predicted or prepared for in advance, attention is assumed to be rapidly updated on-the-fly, triggered by the currently presented context. Context-specific control exemplifies how learning and memory processes can influence attention to enable cognitive flexibility. However, what determines the use of previously learned associations still remains unclear. In the current study, we examined whether task-relevance would influence the learning and use of context cues in a flanker task. Using a secondary counting task, context dimensions associated with differing levels of conflict were made task-relevant or -irrelevant across the experiment. In short, we found that making new contextual information task-relevant caused participants to ignore a previously learned context-attention association and adopt a new context-specific control strategy; all without changing the experimental stimuli. Furthermore, we found participants did not spontaneously learn about context-specific proportion manipulations and explicit instructions insufficient for producing context-specific effects. These results suggests that task-relevance is a key determinant of context-specific control.

@@

Selective attention is commonly investigated using interference paradigms like the Stroop (1935) and flanker (Eriksen \& Eriksen, 1974) tasks, where participants identify a target while ignoring a response-congruent or -incongruent distractor. Performance is typically better on congruent versus incongruent trials and the difference---the congruency effect---taken as an index of attentional priorities. Large congruency effects are thought to reflect ineffective filtering of the distracting stimuli whereas small congruency effects are thought to reflect effective filtering. By probing factors that systematically alter congruency effects, we can then make inferences about processes that control attentional filtering. For example, manipulating the frequency of conflict via the proportion of congruent versus incongruent trials has shown to influence the size of the congruency effect. Typically, a high proportion congruent experiment produces large congruency effects, whereas a low proportion congruent experiment produces small congruency effects (Logan \& Zbrodoff, 1979; Lowe \& Mitterer, 1982; West \& Baylis, 1998). This result is usually explained as strategic control, where participants increase attentional control under high conflict demands and relax attentional control under low conflict demands (Logan, 1980; Logan \& Zbrodoff, 1979; Logan, Zbrodoff, \& Williamson, 1984; Lowe \& Mitterer, 1982). Recent work however, has demonstrated that attentional control is not only adjusted by top-down regulation, but can also be triggered automatically by environmental cues (Brosowsky \& Crump, 2018; Bugg \& Crump, 2012; Egner, 2014; Fischer \& Dreisbach, 2015; King, Korb, \& Egner, 2012; Mayr \& Bryck, 2007).

@@

The current study addresses two related issues: First, how is attentional control resolved when multiple contextual features could be used as cues? And second, what determines the expression and/or suppression of previously learned attentional associations?

Real-world environments are complex and contain many potential cues for directing attention. For instance, a coffee shop is a rich environment with numerous context cues (e.g., pictures on the wall, chairs, coffee, laptop), each of which someone could have different experiences with and afford different learned associations. A laptop for example, could cue attentional processing helpful for writing while other features of the coffee shop might cue attentional processing related to other experiences like socializing and people-watching. Given that all of these cues are available to set attentional priorities, what determines which associative relationships will be used to guide attention? On the one hand, context-based retrieval of attentional priorities might be completely obligatory: the mere presence of learned cues might trigger attention in a ballistic manner (e.g., Schneider \& Shiffrin, 1977; Shiffrin \& Schneider, 1977). Attention in this case would be at the mercy of the environment and attention is guided by the strongest association. On the other hand, our current task-goals might constrain the expression of previously learned associations. If someone sits down to write in the coffee shope, their current goals may allow the laptop to take priority over other environmental cues that are irrelevant to the task at hand.

Historically, perspectives of automaticity and attentional control have emphasized the role of practice and likened the learning of attentional associations to the development of automaticity in skill-learning (Logan, 1988, 1992; Schneider \& Shiffrin, 1977; Shiffrin \& Schneider, 1977). Under this view, forming an association requires many repeated experiences pairing a cue with a set of attentional priorities. Once formed however, it becomes habitual or hard-wired and the mere presence of the cue is sufficient to automatically adjust attention. Learned associations are therefore ballistic, difficult to suppress, and difficult to unlearn (e.g., Schneider \& Shiffrin, 1977; Posner \& Snyder, 1975a, 1975b; Shiffrin \& Schneider, 1977; Treisman, 1960, 1964, 1969). This view is often referenced in modern work on contextual cuing where it is assumed that learned associations require practice (e.g., Lehle \& Hübner, 2008; Crump, 2016), form without awareness or intention, and are triggered automatically by the mere presence of the cue (Crump et al., 2006; Crump \& Logan, 2010; Crump, Vaquero, \& Milliken, 2008; King et al., 2012; Reuss, Desender, Kiesel, \& Kunde, 2014; though, see Brosowsky \& Crump, 2016).

An alternative view---the memory-based perspective---takes a more nuanced view (Abrahamse et al., 2016; Brosowsky \& Crump, 2018; Crump, 2016; Egner, 2014; Norman, 1969, 1976). Here, practice is only important to the extent that it improves memory encoding and retrieval. Learning an association can be thought of as the process of encoding experiences where attentional priorities and cues co-occur. The expression of previously learned associations however, is dictated by memory retrieval---a similarity-based, cue-driven process (e.g., Jacoby, 1978; Hintzman, 1984, 1986, 1988). Presumably, the features of our immediate environment automatically cue memory and reinstate the retrieved attentional priorites (Logan, 1988, 1992). If a few prior experiences are easily retrievable (i.e., memories are distinct or very similar to the retrieval cue), they could support the reinstatement of a learned attentional association despite very little practice (e.g., Brosowsky \& Crump, 2018). Conversely, if prior experiences are difficult to retrieve (i.e., memories are indistinct or low in similarity to the retrieval cue), then learned associations may not be expressed, even after much practice. To the extent that you could manipulate memory retrieval, you should als be able manipulate the expression of learned associations.

Here, we examine whether task-relevance influences the retrieval of previously used attentional priorities. We hypothesized that the cue-driven, memory-retrieval process might be constrained to the features of our environment that are relevant to our on-going task and goals. If this is the case, then a previously learned association should be easily ignored if that feature was made irrelevant to the task because that feature would no longer contribute to memory-retrieval. Furthermore, it would suggest that the mere presence of a cue is not sufficient for automatically directing attention, but instead requires task-integration.

Recent work on contextual cueing suggests that task-relevance is important for establishing new context-attention associations. Crump, Vaquero, \& Milliken (2008) used a Stroop prime-probe task where the distractor word is presented first followed by a color patch probe. Critically, they used the shape of the color patch probe, either a square or a circle, as the contextual dimension; one associated with a high frequency of conflict, the other with a low frequency. In one experiment participants were made aware of the shape-proportion contingencies but given no additional instructions. In the second, participants were given a secondary task to count the number of trials that contained a square. The first experiment failed to find evidence for context-specificity, whereas the second did. Although not testing task-relevance directly, Cañadas, Rodríguez-Bailón, Milliken, and Lupiáñez (2013) found similar effects. In this study they used images of male and female faces as context cues in a flanker task. When given instructions to think of the faces as members of the gender categories they found CSPC effects, but when given instructions to think of the faces as individuals they did not. These results suggest that making a context dimension relevant to the on-going task is important for establishing new associations and producing CSPC effects. However, the extent to which task-relevance influences the expression of previously learned associations is still unknown.

@@

Prior work suggests that the task-relevance of contextual information could be important for establishing new associations between context cues and attentional priorities. Simply put, CSPC effects tend to emerge when context cues are somehow made relevant to the task, but do not when made irrelevant (Cañadas et al., 2013; Crump et al., 2006, 2008). However, what determines the expression of previously learned context-attention associations remains unclear. In the current study, we examined whether manipulating task-relevance would allow participants to ignore a previously learned association and adopt a new context-specific control strategy.

@@

The critical measure of interest is the congruency effect produced in the frequency unbiased context. If, on the one hand, task-relevance has no impact on the use of context-attention associations we would expect no differences between the congruency effects when the high conflict context dimension is made task-relevant or the low conflict context dimension is made task-relevant. This could occur because all three contexts are treated as individual contexts throughout the whole experiment (e.g., Cañadas et al., 2013). Or it could occur because associations formed in the first block interfere with learning new associations in the second (e.g., Brosowsky \& Crump, 2016). On other hand, task-relevance may dictate the expression of previously learned associations. In this case, we would expect smaller congruency effects when the high conflict context dimension is made task-relevant as compared to when the low conflict context dimension is made task-relevant. This would demonstrate that participants were able to ignore a previously learned association and adopt a new context-specific control strategy.

@@

Experiment 1 used a 2x2x3 mixed design with task-relevant context (high conflict and low conflict) and unbiased-item congruency (congruent and incongruent) as within-subject factors, and context-type (object, social, and social/non-repeating) as the between-subjects factor.

@@

\begin{table}[htbp]
\caption{Reaction times and error rates from Experiment 1.}
\label{TR_table}
\centering
\begin{tabular}{lllcccc}
\toprule
 & & & \multicolumn{2}{c}{Congruent} & \multicolumn{2}{c}{Incongruent} \\
\cmidrule(rl){4-5}
\cmidrule(rl){6-7}
\multicolumn{1}{c}{Task-Relevant Context} & \multicolumn{1}{c}{Condition} & \multicolumn{1}{c}{PC} & \multicolumn{1}{c}{RT} & \multicolumn{1}{c}{ER} & \multicolumn{1}{c}{RT} & \multicolumn{1}{c}{ER}  \\
\midrule
\multirow{9}{*}{Low Conflict} & \multirow{3}{*}{Object} & \multicolumn{1}{l}{0\%} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{693 (26)} & \multicolumn{1}{l}{4.17 (0.73)} \\
& & \multicolumn{1}{l}{50\%} & \multicolumn{1}{l}{585 (23)} & \multicolumn{1}{l}{0.74 (0.27)} & \multicolumn{1}{l}{675 (28)} & \multicolumn{1}{l}{4.81 (0.82)} \\
& & \multicolumn{1}{l}{100\%} & \multicolumn{1}{l}{587 (25)} & \multicolumn{1}{l}{0.69 (0.23)} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} \\
\cmidrule(rl){2-7}
& \multirow{3}{*}{Social} & \multicolumn{1}{l}{0\%} & \multicolumn{1}{c}{-} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{745 (25)} & \multicolumn{1}{l}{5.86 (0.77)} \\
& & \multicolumn{1}{l}{50\%} & \multicolumn{1}{l}{612 (21)} & \multicolumn{1}{l}{0.39 (0.19)} & \multicolumn{1}{l}{722 (26)} & \multicolumn{1}{l}{6.78 (1.04)} \\
& & \multicolumn{1}{l}{100\%} & \multicolumn{1}{l}{616 (20)} & \multicolumn{1}{l}{0.63 (0.2)} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} \\
\cmidrule(rl){2-7}
& \multirow{3}{*}{Social (NR)} & \multicolumn{1}{l}{0\%} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{813 (25)} & \multicolumn{1}{l}{5.04 (0.74)} \\
& & \multicolumn{1}{l}{50\%} & \multicolumn{1}{l}{689 (23)} & \multicolumn{1}{l}{0.19 (0.14)} & \multicolumn{1}{l}{793 (27)} & \multicolumn{1}{l}{5.72 (1.06)} \\
& & \multicolumn{1}{l}{100\%} & \multicolumn{1}{l}{689 (25)} & \multicolumn{1}{l}{0.48 (0.17)} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} \\
\midrule
\multirow{9}{*}{High Conflict} & \multirow{3}{*}{Object} & \multicolumn{1}{l}{0\%} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{649 (16)} & \multicolumn{1}{l}{3.84 (0.6)} \\
& & \multicolumn{1}{l}{50\%} & \multicolumn{1}{l}{587 (17)} & \multicolumn{1}{l}{0.56 (0.21)} & \multicolumn{1}{l}{657 (17)} & \multicolumn{1}{l}{4.17 (1.02)} \\
& & \multicolumn{1}{l}{100\%} & \multicolumn{1}{l}{634 (18)} & \multicolumn{1}{l}{2.69 (0.58)} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} \\
\cmidrule(rl){2-7}
& \multirow{3}{*}{Social} & \multicolumn{1}{l}{0\%} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{722 (23)} & \multicolumn{1}{l}{5.77 (0.92)} \\
& & \multicolumn{1}{l}{50\%} & \multicolumn{1}{l}{633 (20)} & \multicolumn{1}{l}{0.78 (0.5)} & \multicolumn{1}{l}{722 (25)} & \multicolumn{1}{l}{4.75 (0.9)} \\
& & \multicolumn{1}{l}{100\%} & \multicolumn{1}{l}{673 (21)} & \multicolumn{1}{l}{1.84 (0.47)} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} \\
\cmidrule(rl){2-7}
& \multirow{3}{*}{Social (NR)} & \multicolumn{1}{l}{0\%} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{807 (26)} & \multicolumn{1}{l}{4.55 (0.8)} \\
& & \multicolumn{1}{l}{50\%} & \multicolumn{1}{l}{735 (28)} & \multicolumn{1}{l}{0.87 (0.36)} & \multicolumn{1}{l}{818 (30)} & \multicolumn{1}{l}{4.26 (0.8)} \\
& & \multicolumn{1}{l}{100\%} & \multicolumn{1}{l}{779 (34)} & \multicolumn{1}{l}{2.03 (0.4)} & \multicolumn{1}{l}{-} & \multicolumn{1}{l}{-} \\
 & & & & & & \\
\bottomrule
\multicolumn{7}{l}{\textit{Note}: RT = Reaction Time (ms);  ER = Error Rates (\%); PC = Proportion Congruent;} \\
\multicolumn{7}{l}{NR = Non-Repeating; Standard Errors are presented in parantheses.} \\
\end{tabular}%
\end{table}

@@

@@

In Experiment 1, we found that manipulating the task-relevance of a shared contextual feature was associated in predictable changes in context-specific proportion congruency effects for frequency unbiased context items. Namely, we found larger congruency effects when the context was associated with low conflict than when the context was associated with high conflict. This suggests that participants learned the association between context and conflict-level in the first block (despite multiple conetxt cues), and subsequently ignored that association in the second block to learn a new association. Task-relevance then, may not just be important for establishing new associations (e.g., Crump \& Milliken, 2009) but also important for determining which learned associations are expressed in the present moment.

@@

Experiment 2, therefore, serves as a control experiment to test these two assumptions directly. Experiment 2 is identical to Experiment 1 with two important changes: First, we removed the counting task, giving no explicit instructions about how to use the context cues. We only informed participants that images would be present throughout the task. This allowed us to determine if there was any systematic effect of biased items on the unbiased items absent the counting task. Second, we altered the proportion congruency of the biased item sets from 100\%/0\% to 75\%/25\%. This allowed us to measure the congruency effect for the biased item sets to determine whether participants learned about the associations between proportion congruency and the context cues without the counting task (e.g., Crump \& Milliken, 2009). If this were the case, we would expect larger congruency effects for the 75\% proportion congruent items as compared to the 25\% proportion congruent items.

@@

\begin{table}[htbp]
\caption{Reaction times and error rates from Experiment 2.}
\label{TR_table_2}
\centering
\begin{tabular}{lcccc}
\toprule
& \multicolumn{2}{c}{Congruent} & \multicolumn{2}{c}{Incongruent} \\
\cmidrule(rl){2-3}
\cmidrule(rl){4-5}
\multicolumn{1}{c}{PC} & \multicolumn{1}{c}{RT} & \multicolumn{1}{c}{ER} & \multicolumn{1}{c}{RT} & \multicolumn{1}{c}{ER}  \\
\midrule
\multicolumn{1}{l}{25\%} & \multicolumn{1}{l}{518 (16)} & \multicolumn{1}{l}{0.35 (0.17)} & \multicolumn{1}{l}{598 (17)} & \multicolumn{1}{l}{3.78 (0.71)} \\
\multicolumn{1}{l}{50\%} & \multicolumn{1}{l}{515 (15)} & \multicolumn{1}{l}{0.27 (0.1)} & \multicolumn{1}{l}{603 (17)} & \multicolumn{1}{l}{4.26 (0.7)} \\
\multicolumn{1}{l}{100\%} & \multicolumn{1}{l}{523 (18)} & \multicolumn{1}{l}{0.18 (0.07)} & \multicolumn{1}{l}{603 (19)} & \multicolumn{1}{l}{3.72 (0.61)} \\
\bottomrule
\multicolumn{5}{l}{\textit{Note}: RT = Reaction Time (ms);  ER = Error Rates (\%);} \\
\multicolumn{5}{l}{PC = Proportion Congruent; Standard Errors} \\
\multicolumn{5}{l}{are presented in parantheses.} \\
\end{tabular}%
\end{table}

@@

The goal of the current study has been to examine the importance of task-relevance for the contextual recruitment of selective attention. In Experiment 1, we found that task-relevance allowed participants to exploit context cues in the presence of a competing, contradictory cue and subsequently ignore a previously learned association in favor of a different context cue. In Experiment 2, we conceptually replicated previous work showing that task-relevance was necessary for participants to learn about the biased proportion congruency associations (e.g., Crump \& Milliken, 2009) and ruled out possible confounding effects of the biased items.

Experiment 3 tests the boundaries of the task-relevance effect found in Experiment 1. The evidence thus far suggests that context-cues, rather than being irrelevant to the task (as their namesake would suggest), need to be integrated into the task-representation to guide attentional control. There is some evidence that instructions alone can be sufficient to produce list-wide proportion congruency effects, though not item-specific proportion congruency effects (Desender, 2018; Entel, Tzelgov, \& Bereby-Meyer, 2014) and not context-specific proportion congruency effects (Crump \& Milliken, 2009). Explicit instructions however, are another way of making context-cues relevant to the on-going task and could serve the same purpose as the counting task. In Experiment 3, we test whether participants are able to explicitly use the context-cues to guide attentional control through an instruction manipulation.

@@

\newpage

@@

\leavevmode\hypertarget{ref-open_science_collaboration_estimating_2015}{}%
Collaboration, O. S. (2015). Estimating the reproducibility of psychological science. \emph{Science}, \emph{349}(6251), aac4716. doi:\href{https://doi.org/10.1126/science.aac4716}{10.1126/science.aac4716}
